### 2025-04-09 paper/paper.qmd
Outline

---
title: "STA107 Post-Course Survey Analysis: Evaluating Student Reflections on R and Statistical Learning"
author: 
  - Sinan Ma
  - Jaiditya Dev
thanks: "Thanks to the STA107 students for participating in this anonymous post-course survey and for sharing their feedback."
date: today
date-format: long
abstract: "This report analyzes feedback collected from a post-course survey in STA107 (Introduction to Statistics), focusing on students’ experiences with R, the integration of normal models, and their general reflections on the course structure. Using both quantitative and qualitative methods, we identify key trends and themes in students’ learning experiences to inform future curriculum design and pedagogical improvements."
format: pdf
toc: true
number-sections: true
bibliography: references.bib
pdf-engine: xelatex
---

```{r}
#| include: false
#| warning: false
#| message: false

#### workspace setup ####
library(tidyverse)
library(readr)
library(janitor)
library(ggplot2)
```

```{r}
#| include: false

#### import survey data ####
survey_data <- read_csv(here("data/00-raw_data.csv")) 
# change to clean data later
```

# Introduction

This report presents an analysis of the anonymous post-course survey completed by students enrolled in **STA107: Introduction to Statistics** at the University of Toronto Mississauga. The survey was voluntary, open for multiple submissions, and contributed **2% to the course grade**. Its goal was to evaluate students’ experiences with R-based activities, the integration of normal distribution concepts, and overall satisfaction with the course components. This feedback will help refine future course offerings and improve the learning experience.

# Survey Questions

This section outlines the structure of the survey and highlights key question types:

- **Quantitative items**: Likert-scale responses on course clarity, usefulness of R exercises, and understanding of statistical concepts.
- **Qualitative items**: Open-ended prompts for students to share what they found most/least helpful and suggestions for improvement.

# Data Analysis

## Methodology

We employed a **mixed-methods approach**:

- **Quantitative analysis**: Descriptive statistics and visualizations for Likert-scale questions.
- **Qualitative analysis**: Thematic coding of open-ended responses to identify common sentiments and suggestions.

# Results

## Quantitative Results

- Distribution of responses to Likert-scale questions
- Summary statistics (mean, median, SD)
- Visualizations: bar plots, histograms

## Qualitative Results

- Key themes identified
- Word frequency or sentiment analysis (optional)

# Discussion

- **Interpretation** of major findings: strengths and gaps in course delivery
- **Limitations**: sampling bias, multiple submissions allowed, missing data
- **Future directions**: refining R activities, scaffolding normal model lessons, improving student support

# Conclusion

The STA107 post-course survey provides valuable insights into how students perceive the course structure and its technical components. Findings highlight both effective practices and opportunities for enhancement. Continued evaluation and iteration will ensure the course remains responsive to student needs and pedagogical best practices.

# References

Citations goes here.

### 2025-04-09 scripts/00-data_cleaning.R
cleaned data, and added scripts for it

@@ -1,0 +1,27 @@
+#### Preamble ####
+# Purpose: Cleaning STA107 post-course survey data
+# Author: Sinan Ma and Jaiditya Dev
+# Email: sinan.ma@mail.utoronto.ca
+# Date: 9 April 2025
+
+#### Workspace setup ####
+library(tidyverse)
+library(janitor)
+
+#### Load raw data ####
+survey_raw <- read_csv("data/raw_data.csv")
+
+#### Clean data ####
+survey_clean <- survey_raw |> 
+  clean_names() |> 
+  select(
+    -section, -section_id, -section_sis_id, -submitted, -attempt,
+    -n_correct, -n_incorrect, -score,
+    -starts_with("x1_")
+  )
+
+#### Save cleaned data ####
+write_csv(
+  x = survey_clean,
+  file = "data/clean_data.csv"
+)

### 2025-04-09 scripts/02-qualitative_data_filtering.R

@@ -1,0 +1,28 @@
+#### Preamble ####
+# Purpose: Filtering qualitative data from STA107 post-course survey
+# Author: Sinan Ma and Jaiditya Dev
+# Email: sinan.ma@mail.utoronto.ca
+# Date: 9 April 2025
+
+#### Workspace setup ####
+library(tidyverse)
+library(janitor)
+
+#### Load cleaned data ####
+survey_clean <- read_csv("data/clean_data.csv")
+
+#### Filter qualitative data ####
+qual_data <- survey_clean |> 
+  select(-starts_with("x4911959"),
+         -starts_with("x4911960"),
+         -starts_with("x4911962"),
+         -starts_with("x4911963"),
+         -starts_with("x4911996"),
+         -starts_with("x4911965")) |> 
+  filter(if_any(everything(), ~ !is.na(.)))
+
+#### Save filtered data ####
+write_csv(
+  x = qual_data,
+  file = "data/qualitative_analysis_data.csv"
+)

### 2025-04-09 scripts/01-quantitative_data_filtering.R
cleaned data, and added scripts for it

@@ -1,0 +1,27 @@
+#### Preamble ####
+# Purpose: Filtering quantitative data from STA107 post-course survey
+# Author: Sinan Ma and Jaiditya Dev
+# Email: sinan.ma@mail.utoronto.ca
+# Date: 9 April 2025
+
+#### Workspace setup ####
+library(tidyverse)
+library(janitor)
+
+#### Load cleaned data ####
+survey_clean <- read_csv("data/clean_data.csv")
+
+#### Filter quantitative data ####
+quant_data <- survey_clean |> 
+  select(starts_with("x4911959"),
+         starts_with("x4911960"),
+         starts_with("x4911962"),
+         starts_with("x4911963"),
+         starts_with("x4911996"),
+         starts_with("x4911965"))
+
+#### Save filtered data ####
+write_csv(
+  x = quant_data,
+  file = "data/quantitative_analysis_data.csv"
+)

### 2025-04-13 scripts/03-analysis_qualitative.R
Added Qualitative Analysis

@@ -1,0 +1,71 @@
+# Load required libraries
+library(tidyverse)         # For data manipulation (includes dplyr, ggplot2, etc.)
+library(stringr)           # For string handling like trimming, pattern matching
+library(tidytext)          # For text mining and tokenization
+library(syuzhet)           # For sentiment analysis using NRC lexicon
+library(ggplot2)           # For creating visualizations (e.g., bar plots)
+library(tidyr)             # For data reshaping (e.g., pivot_longer)
+library(wordcloud)         # For creating word cloud visualizations
+
+# Read in the survey data
+data <- read.csv("data/qualitative_analysis_data.csv", stringsAsFactors = FALSE)
+
+# Rename columns
+colnames(data)[1:11] <- c("Q3", "Q6", "Q8", "Q10", "Q11", "Q12", "Q13", "Q14", "Q15", "Q16", "Q17")
+
+# Clean the text
+data_clean <- data %>%
+  mutate(across(everything(), ~ str_trim(tolower(as.character(.)))))
+
+# Initialize an empty data frame to collect sentiment results
+sentiment_results <- data.frame(
+  Question = character(),
+  Positive = numeric(),
+  Negative = numeric(),
+  Neutral = numeric(),
+  stringsAsFactors = FALSE
+)
+
+# Define function to calculate sentiment per question
+analyze_sentiment <- function(question_col, question_name) {
+  sentiments <- get_nrc_sentiment(data_clean[[question_col]])
+  
+  # Calculate positive and negative
+  pos <- sum(sentiments$positive)
+  neg <- sum(sentiments$negative)
+  
+  # Define neutral as total rows - (pos > 0 or neg > 0)
+  neutral <- nrow(sentiments) - sum(rowSums(sentiments[, c("positive", "negative")]) > 0)
+  
+  # Append to results
+  sentiment_results <<- rbind(sentiment_results, data.frame(
+    Question = question_name,
+    Positive = pos,
+    Negative = neg,
+    Neutral = neutral
+  ))
+}
+
+# List of questions to analyze
+question_columns <- c("Q3", "Q6", "Q8", "Q10", "Q11", "Q12", "Q13", "Q14", "Q15", "Q16", "Q17")
+
+# Apply sentiment analysis
+walk2(question_columns, question_columns, analyze_sentiment)
+
+# View results
+print(sentiment_results)
+
+# Reshape sentiment_results for plotting
+sentiment_long <- sentiment_results %>%
[…]

### 2025-04-14 scripts/03-analysis_qualitative.R
Fixed Graph Qual. Analysis

@@ -46,22 +46,13 @@ analyze_sentiment <- function(question_col, question_name) {
   ))
 }
 
-# List of questions to analyze
-question_columns <- c("Q3", "Q6", "Q8", "Q10", "Q11", "Q12", "Q13", "Q14", "Q15", "Q16", "Q17")
+# Define the desired order of questions
+ordered_questions <- c("Q3", "Q6", "Q8", "Q10", "Q11", "Q12", "Q13", "Q14", "Q15", "Q16", "Q17")
 
-# Apply sentiment analysis
-walk2(question_columns, question_columns, analyze_sentiment)
+# Convert the 'Question' column to a factor with specified levels
+sentiment_long$Question <- factor(sentiment_long$Question, levels = ordered_questions)
 
-# View results
-print(sentiment_results)
-
-# Reshape sentiment_results for plotting
-sentiment_long <- sentiment_results %>%
-  pivot_longer(cols = c("Positive", "Negative", "Neutral"),
-               names_to = "Sentiment",
-               values_to = "Count")
-
-# Plot
+# Plot with corrected order
 ggplot(sentiment_long, aes(x = Question, y = Count, fill = Sentiment)) +
   geom_bar(stat = "identity", position = "dodge") +
   scale_fill_manual(values = c("Positive" = "darkgreen", "Negative" = "red", "Neutral" = "gray")) +
@@ -69,3 +60,4 @@ ggplot(sentiment_long, aes(x = Question, y = Count, fill = Sentiment)) +
   labs(title = "Sentiment Distribution by Survey Question",
        x = "Survey Question",
        y = "Sentiment Count")
+

### 2025-04-14 paper/paper.qmd
added paper.qmd

@@
 ---
-title: "STA107 Post-Course Survey Analysis: Evaluating Student Reflections on R and Statistical Learning"
+title: "STA107 Post-Course Survey Analysis"
+subtitle: "Evaluating Student Reflections on R and Statistical Learning"
 author: 
@@
   - Jaiditya Dev
-thanks: "Thanks to the STA107 students for participating in this anonymous post-course survey and for sharing their feedback."
-date: today
-date-format: long
-abstract: "This report analyzes feedback collected from a post-course survey in STA107 (Introduction to Statistics), focusing on students’ experiences with R, the integration of normal models, and their general reflections on the course structure.
-Using both quantitative and qualitative methods, we identify key trends and themes in students’ learning experiences to inform future curriculum design and pedagogical improvements."
-format: pdf
+date: "April 13, 2025"
+format: 
+  pdf:
     toc: true
     number-sections: true
-bibliography: references.bib
-pdf-engine: xelatex
+    code-fold: true
+    code-summary: "Show Code"
+    fig-align: center
+    fig-cap-location: top
+    keep-tex: true
 ---
@@
 library(janitor)
+library(here)
 library(ggplot2)
@@
 
-```{r}
-#| include: false
+```{r setup, include=FALSE}
 
 #### import survey data ####
-survey_data <- read_csv(here("data/00-raw_data.csv")) 
+data <- read_csv("../data/clean_data.csv")
+qualitative_data <- read_csv("../data/qualitative_analysis_data.csv")
 # change to clean data later
@@
 
+The post-course survey comprised 17 open- and close-ended questions.
+Among these, 11 were qualitative in nature (`Q3`, `Q6`, `Q8`, `Q10`, `Q11`, `Q12`, `Q13`, `Q14`, `Q15`, `Q16`, `Q17`), focusing on free-text responses reflecting students' thoughts on the content, instruction, and overall experience.
+
 # Data Analysis
@@
 - **Qualitative analysis**: Thematic coding of open-ended responses to identify common sentiments and suggestions.
+
+### Qualitative Analysis
+```{r}
+# Load required libraries
+library(tidyverse)
+library(stringr)
+library(tidytext)
+library(syuzhet)
+library(ggplot2)
+library(tidyr)
+library(wordcloud)
+
+# Read in the survey data
+data <- qualitative_data
+
+# Rename columns
+colnames(data)[1:11] <- c("Q3", "Q6", "Q8", "Q10", "Q11", 
+                          "Q12", "Q13", "Q14", "Q15", "Q16", "Q17")
+
+# Clean the text
+data_clean <- data %>%
+  mutate(across(everything(), ~ str_trim(tolower(as.character(.)))))
+
+# Initialize an empty data frame to collect sentiment results
+sentiment_results <- data.frame(
+  Question = character(),
+  Positive = numeric(),
+  Negative = numeric(),
+  Neutral = numeric(),
+  stringsAsFactors = FALSE
+)
+
+# Function to calculate sentiment
+analyze_sentiment <- function(question_col, question_name) {
+  sentiments <- get_nrc_sentiment(data_clean[[question_col]])
+
+  pos <- sum(sentiments$positive)
+  neg <- sum(sentiments$negative)
+  neutral <- nrow(sentiments) - 
+    sum(rowSums(sentiments[, c("positive", "negative")]) > 0)
+
+  sentiment_results <<- rbind(sentiment_results, data.frame(
+    Question = question_name,
+    Positive = pos,
+    Negative = neg,
+    Neutral = neutral
+  ))
+}
+
+# Apply the function to all relevant questions
+question_names <- c("Q3", "Q6", "Q8", "Q10", "Q11", 
+                    "Q12", "Q13", "Q14", "Q15", "Q16", "Q17")
+
+for (qn in question_names) {
+  analyze_sentiment(qn, qn)
+}
+```
 
@@
 
-- Key themes identified
-- Word frequency or sentiment analysis (optional)
+The sentiment analysis revealed a clear trend of positive student engagement across most open-ended responses.
+When asked about what aspects of the course were most helpful, students consistently expressed appreciation for the structured content delivery, use of R, and clarity in statistical explanations.
+Similarly, reflections on the usefulness of the normal model and applied examples elicited strong positive sentiment, highlighting the effectiveness of concept integration throughout the term.
 
-# Discussion
+Feedback on the clarity of teaching and accessibility of course material was more mixed.
+While many students praised the instructor’s effort and teaching style, others noted inconsistencies in understanding due to the pace of delivery or difficulty following complex R assignments.
 
-- **Interpretation** of major findings: strengths and gaps in course delivery
-- **Limitations**: sampling bias, multiple submissions allowed, missing data
-- **Future directions**: refining R activities, scaffolding normal model lessons, improving student support
+The most critical responses were seen in prompts asking what aspects of the course were least helpful and what improvements could be made.
+Here, students frequently cited the fast pace of certain topics, confusion around R syntax, and a need for more real-world or scaffolded examples.
+Words like “confusing,” “fast,” and “examples” frequently appeared in these responses, pointing to the need for clearer instructional supports.
+
+```{r sentiment-plot, echo=FALSE, fig.cap="Sentiment Distribution by Survey Question"}
+# Reshape the data to long format for plotting
+sentiment_long <- sentiment_results %>%
+  pivot_longer(cols = c("Positive", "Negative", "Neutral"), 
+               names_to = "Sentiment", 
+               values_to = "Count")
+
+# Set correct order for x-axis
+sentiment_long$Question <- factor(sentiment_long$Question,
+                                  levels = question_names)
+
+# Plot
+ggplot(sentiment_long, aes(x = Question, y = Count, fill = Sentiment)) +
+  geom_bar(stat = "identity", position = "dodge") +
+  scale_fill_manual(values = c("Positive" = "darkgreen", 
+                               "Negative" = "red", "Neutral" = "gray")) +
+  theme_minimal() +
+  labs(title = "Sentiment Distribution by Survey Question",
+       x = "Survey Question",
+       y = "Sentiment Count")
+```
 
@@
 
+From a qualitative perspective, student feedback revealed strong appreciation for several key elements of the course—especially the hands-on R activities, the integration of real-world statistical examples, and the clarity of normal model explanations.
+These components were described as engaging and instructive, affirming their value in helping students apply theoretical knowledge in practical settings.
+
+However, when students were asked about areas of improvement, many raised concerns regarding the speed at which complex topics were introduced, as well as the difficulty of interpreting R-based assignments without more structured guidance.
+Themes of pacing, clarity, and the need for more scaffolded instruction were particularly prominent.
+Addressing these concerns by slowing down the delivery of technical material, providing more detailed examples, and incorporating mid-semester feedback opportunities could significantly enhance the learning experience in future offerings of STA107.
+
 # References
 
-Citations goes here.
+- Jockers, M. (2017). *syuzhet: Extracts Sentiment and Sentiment-Derived Plot Arcs from Text* (R package v1.0.4). [CRAN Documentation](https://cran.r-project.org/web/packages/syuzhet/index.html)
+
+- Jockers, M. (2020). *Introduction to the Syuzhet Package*. [CRAN Vignette](https://cran.r-project.org/web/packages/syuzhet/vignettes/syuzhet-vignette.html)
+
+- Kim, H.
+(2022).
+Sentiment Analysis: Limits and Progress of the Syuzhet Package and Its Lexicons.
+*Digital Humanities Quarterly, 16*(2).
+[Article](http://www.digitalhumanities.org/dhq/vol/16/2/000601/000601.html)

### 2025-04-14 paper/paper.tex
added paper.qmd

% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
\PassOptionsToPackage{dvipsnames,svgnames,x11names}{xcolor}
%
\documentclass[
  letterpaper,
  DIV=11,
  numbers=noendperiod]{scrartcl}

\usepackage{amsmath,amssymb}
\usepackage{iftex}
\ifPDFTeX
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math}
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
\usepackage{lmodern}
\ifPDFTeX\else  
    % xetex/luatex font selection
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\setlength{\emergencystretch}{3em} % prevent overfull lines
\setcounter{secnumdepth}{5}
% Make \paragraph and \subparagraph free-standing
\makeatletter
\ifx\paragraph\undefined\else
  \let\oldparagraph\paragraph
  \renewcommand{\paragraph}{
    \@ifstar
      \xxxParagraphStar
      \xxxParagraphNoStar
  }
  \newcommand{\xxxParagraphStar}[1]{\oldparagraph*{#1}\mbox{}}
  \newcommand{\xxxParagraphNoStar}[1]{\oldparagraph{#1}\mbox{}}
\fi
\ifx\subparagraph\undefined\else
  \let\oldsubparagraph\subparagraph
  \renewcommand{\subparagraph}{
    \@ifstar
      \xxxSubParagraphStar
      \xxxSubParagraphNoStar
  }
  \newcommand{\xxxSubParagraphStar}[1]{\oldsubparagraph*{#1}\mbox{}}
  \newcommand{\xxxSubParagraphNoStar}[1]{\oldsubparagraph{#1}\mbox{}}
\fi
\makeatother

\usepackage{color}
\usepackage{fancyvrb}
\newcommand{\VerbBar}{|}
\newcommand{\VERB}{\Verb[commandchars=\\\{\}]}
\DefineVerbatimEnvironment{Highlighting}{Verbatim}{commandchars=\\\{\}}
% Add ',fontsize=\small' for more characters per line
\usepackage{framed}
\definecolor{shadecolor}{RGB}{241,243,245}
\newenvironment{Shaded}{\begin{snugshade}}{\end{snugshade}}
\newcommand{\AlertTok}[1]{\textcolor[rgb]{0.68,0.00,0.00}{#1}}
\newcommand{\AnnotationTok}[1]{\textcolor[rgb]{0.37,0.37,0.37}{#1}}
\newcommand{\AttributeTok}[1]{\textcolor[rgb]{0.40,0.45,0.13}{#1}}
\newcommand{\BaseNTok}[1]{\textcolor[rgb]{0.68,0.00,0.00}{#1}}
\newcommand{\BuiltInTok}[1]{\textcolor[rgb]{0.00,0.23,0.31}{#1}}
\newcommand{\CharTok}[1]{\textcolor[rgb]{0.13,0.47,0.30}{#1}}
\newcommand{\CommentTok}[1]{\textcolor[rgb]{0.37,0.37,0.37}{#1}}
\newcommand{\CommentVarTok}[1]{\textcolor[rgb]{0.37,0.37,0.37}{\textit{#1}}}
\newcommand{\ConstantTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{#1}}
\newcommand{\ControlFlowTok}[1]{\textcolor[rgb]{0.00,0.23,0.31}{\textbf{#1}}}
\newcommand{\DataTypeTok}[1]{\textcolor[rgb]{0.68,0.00,0.00}{#1}}
\newcommand{\DecValTok}[1]{\textcolor[rgb]{0.68,0.00,0.00}{#1}}
\newcommand{\DocumentationTok}[1]{\textcolor[rgb]{0.37,0.37,0.37}{\textit{#1}}}
\newcommand{\ErrorTok}[1]{\textcolor[rgb]{0.68,0.00,0.00}{#1}}
\newcommand{\ExtensionTok}[1]{\textcolor[rgb]{0.00,0.23,0.31}{#1}}
\newcommand{\FloatTok}[1]{\textcolor[rgb]{0.68,0.00,0.00}{#1}}
\newcommand{\FunctionTok}[1]{\textcolor[rgb]{0.28,0.35,0.67}{#1}}
\newcommand{\ImportTok}[1]{\textcolor[rgb]{0.00,0.46,0.62}{#1}}
\newcommand{\InformationTok}[1]{\textcolor[rgb]{0.37,0.37,0.37}{#1}}
\newcommand{\KeywordTok}[1]{\textcolor[rgb]{0.00,0.23,0.31}{\textbf{#1}}}
\newcommand{\NormalTok}[1]{\textcolor[rgb]{0.00,0.23,0.31}{#1}}
\newcommand{\OperatorTok}[1]{\textcolor[rgb]{0.37,0.37,0.37}{#1}}
\newcommand{\OtherTok}[1]{\textcolor[rgb]{0.00,0.23,0.31}{#1}}
\newcommand{\PreprocessorTok}[1]{\textcolor[rgb]{0.68,0.00,0.00}{#1}}
\newcommand{\RegionMarkerTok}[1]{\textcolor[rgb]{0.00,0.23,0.31}{#1}}
\newcommand{\SpecialCharTok}[1]{\textcolor[rgb]{0.37,0.37,0.37}{#1}}
\newcommand{\SpecialStringTok}[1]{\textcolor[rgb]{0.13,0.47,0.30}{#1}}
\newcommand{\StringTok}[1]{\textcolor[rgb]{0.13,0.47,0.30}{#1}}
\newcommand{\VariableTok}[1]{\textcolor[rgb]{0.07,0.07,0.07}{#1}}
\newcommand{\VerbatimStringTok}[1]{\textcolor[rgb]{0.13,0.47,0.30}{#1}}
\newcommand{\WarningTok}[1]{\textcolor[rgb]{0.37,0.37,0.37}{\textit{#1}}}

\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}\usepackage{longtable,booktabs,array}
\usepackage{calc} % for calculating minipage widths
% Correct order of tables after \paragraph or \subparagraph
\usepackage{etoolbox}
\makeatletter
\patchcmd\longtable{\par}{\if@noskipsec\mbox{}\fi\par}{}{}
\makeatother
% Allow footnotes in longtable head/foot
\IfFileExists{footnotehyper.sty}{\usepackage{footnotehyper}}{\usepackage{footnote}}
\makesavenoteenv{longtable}
\usepackage{graphicx}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
% Set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother

\KOMAoption{captions}{tableheading,figureheading}
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\AtBeginDocument{%
\ifdefined\contentsname
  \renewcommand*\contentsname{Table of contents}
\else
  \newcommand\contentsname{Table of contents}
\fi
\ifdefined\listfigurename
  \renewcommand*\listfigurename{List of Figures}
\else
  \newcommand\listfigurename{List of Figures}
\fi
\ifdefined\listtablename
  \renewcommand*\listtablename{List of Tables}
\else
  \newcommand\listtablename{List of Tables}
\fi
\ifdefined\figurename
  \renewcommand*\figurename{Figure}
\else
  \newcommand\figurename{Figure}
\fi
\ifdefined\tablename
  \renewcommand*\tablename{Table}
\else
  \newcommand\tablename{Table}
\fi
}
\@ifpackageloaded{float}{}{\usepackage{float}}
\floatstyle{ruled}
\@ifundefined{c@chapter}{\newfloat{codelisting}{h}{lop}}{\newfloat{codelisting}{h}{lop}[chapter]}
\floatname{codelisting}{Listing}
\newcommand*\listoflistings{\listof{codelisting}{List of Listings}}
\makeatother
\makeatletter
\makeatother
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\@ifpackageloaded{subcaption}{}{\usepackage{subcaption}}
\makeatother

\ifLuaTeX
  \usepackage{selnolig}  % disable illegal ligatures
\fi
\usepackage{bookmark}

\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\urlstyle{same} % disable monospaced font for URLs
\hypersetup{
  pdftitle={STA107 Post-Course Survey Analysis},
  pdfauthor={Sinan Ma; Jaiditya Dev},
  colorlinks=true,
  linkcolor={blue},
  filecolor={Maroon},
  citecolor={Blue},
  urlcolor={Blue},
  pdfcreator={LaTeX via pandoc}}

\title{STA107 Post-Course Survey Analysis}
\usepackage{etoolbox}
\makeatletter
\providecommand{\subtitle}[1]{% add subtitle to \maketitle
  \apptocmd{\@title}{\par {\large #1 \par}}{}{}
}
\makeatother
\subtitle{Evaluating Student Reflections on R and Statistical Learning}
\author{Sinan Ma \and Jaiditya Dev}
\date{2025-04-13}

\begin{document}
\maketitle

\renewcommand*\contentsname{Table of contents}
{
\hypersetup{linkcolor=}
\setcounter{tocdepth}{3}
\tableofcontents
}

\section{Introduction}\label{introduction}

This report presents an analysis of the anonymous post-course survey
completed by students enrolled in \textbf{STA107: Introduction to
Statistics} at the University of Toronto Mississauga. The survey was
voluntary, open for multiple submissions, and contributed \textbf{2\% to
the course grade}. Its goal was to evaluate students' experiences with
R-based activities, the integration of normal distribution concepts, and
overall satisfaction with the course components. This feedback will help
refine future course offerings and improve the learning experience.

\section{Survey Questions}\label{survey-questions}

This section outlines the structure of the survey and highlights key
question types:

\begin{itemize}
\tightlist
\item
  \textbf{Quantitative items}: Likert-scale responses on course clarity,
  usefulness of R exercises, and understanding of statistical concepts.
\item
  \textbf{Qualitative items}: Open-ended prompts for students to share
  what they found most/least helpful and suggestions for improvement.
\end{itemize}

The post-course survey comprised 17 open- and close-ended questions.
Among these, 11 were qualitative in nature (\texttt{Q3}, \texttt{Q6},
\texttt{Q8}, \texttt{Q10}, \texttt{Q11}, \texttt{Q12}, \texttt{Q13},
\texttt{Q14}, \texttt{Q15}, \texttt{Q16}, \texttt{Q17}), focusing on
free-text responses reflecting students' thoughts on the content,
instruction, and overall experience.

\section{Data Analysis}\label{data-analysis}

\subsection{Methodology}\label{methodology}

We employed a \textbf{mixed-methods approach}:

\begin{itemize}
\tightlist
\item
  \textbf{Quantitative analysis}: Descriptive statistics and
  visualizations for Likert-scale questions.
\item
  \textbf{Qualitative analysis}: Thematic coding of open-ended responses
  to identify common sentiments and suggestions.
\end{itemize}

\subsubsection{Qualitative Analysis}\label{qualitative-analysis}

\begin{Shaded}
\begin{Highlighting}[]
\CommentTok{\# Load required libraries}
\FunctionTok{library}\NormalTok{(tidyverse)}
\FunctionTok{library}\NormalTok{(stringr)}
\FunctionTok{library}\NormalTok{(tidytext)}
\FunctionTok{library}\NormalTok{(syuzhet)}
\FunctionTok{library}\NormalTok{(ggplot2)}
\FunctionTok{library}\NormalTok{(tidyr)}
\FunctionTok{library}\NormalTok{(wordcloud)}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
Loading required package: RColorBrewer
\end{verbatim}

\begin{Shaded}
\begin{Highlighting}[]
\CommentTok{\# Read in the survey data}
\NormalTok{data }\OtherTok{\textless{}{-}}\NormalTok{ qualitative\_data}

\CommentTok{\# Rename columns}
\FunctionTok{colnames}\NormalTok{(data)[}\DecValTok{1}\SpecialCharTok{:}\DecValTok{11}\NormalTok{] }\OtherTok{\textless{}{-}} \FunctionTok{c}\NormalTok{(}\StringTok{"Q3"}\NormalTok{, }\StringTok{"Q6"}\NormalTok{, }\StringTok{"Q8"}\NormalTok{, }\StringTok{"Q10"}\NormalTok{, }\StringTok{"Q11"}\NormalTok{, }
                          \StringTok{"Q12"}\NormalTok{, }\StringTok{"Q13"}\NormalTok{, }\StringTok{"Q14"}\NormalTok{, }\StringTok{"Q15"}\NormalTok{, }\StringTok{"Q16"}\NormalTok{, }\StringTok{"Q17"}\NormalTok{)}

\CommentTok{\# Clean the text}
\NormalTok{data\_clean }\OtherTok{\textless{}{-}}\NormalTok{ data }\SpecialCharTok{\%\textgreater{}\%}
  \FunctionTok{mutate}\NormalTok{(}\FunctionTok{across}\NormalTok{(}\FunctionTok{everything}\NormalTok{(), }\SpecialCharTok{\textasciitilde{}} \FunctionTok{str\_trim}\NormalTok{(}\FunctionTok{tolower}\NormalTok{(}\FunctionTok{as.character}\NormalTok{(.)))))}

\CommentTok{\# Initialize an empty data frame to collect sentiment results}
\NormalTok{sentiment\_results }\OtherTok{\textless{}{-}} \FunctionTok{data.frame}\NormalTok{(}
  \AttributeTok{Question =} \FunctionTok{character}\NormalTok{(),}
  \AttributeTok{Positive =} \FunctionTok{numeric}\NormalTok{(),}
  \AttributeTok{Negative =} \FunctionTok{numeric}\NormalTok{(),}
  \AttributeTok{Neutral =} \FunctionTok{numeric}\NormalTok{(),}
  \AttributeTok{stringsAsFactors =} \ConstantTok{FALSE}
\NormalTok{)}

\CommentTok{\# Function to calculate sentiment}
\NormalTok{analyze\_sentiment }\OtherTok{\textless{}{-}} \ControlFlowTok{function}\NormalTok{(question\_col, question\_name) \{}
\NormalTok{  sentiments }\OtherTok{\textless{}{-}} \FunctionTok{get\_nrc\_sentiment}\NormalTok{(data\_clean[[question\_col]])}

\NormalTok{  pos }\OtherTok{\textless{}{-}} \FunctionTok{sum}\NormalTok{(sentiments}\SpecialCharTok{$}\NormalTok{positive)}
\NormalTok{  neg }\OtherTok{\textless{}{-}} \FunctionTok{sum}\NormalTok{(sentiments}\SpecialCharTok{$}\NormalTok{negative)}
\NormalTok{  neutral }\OtherTok{\textless{}{-}} \FunctionTok{nrow}\NormalTok{(sentiments) }\SpecialCharTok{{-}} 
    \FunctionTok{sum}\NormalTok{(}\FunctionTok{rowSums}\NormalTok{(sentiments[, }\FunctionTok{c}\NormalTok{(}\StringTok{"positive"}\NormalTok{, }\StringTok{"negative"}\NormalTok{)]) }\SpecialCharTok{\textgreater{}} \DecValTok{0}\NormalTok{)}

\NormalTok{  sentiment\_results }\OtherTok{\textless{}\textless{}{-}} \FunctionTok{rbind}\NormalTok{(sentiment\_results, }\FunctionTok{data.frame}\NormalTok{(}
    \AttributeTok{Question =}\NormalTok{ question\_name,}
    \AttributeTok{Positive =}\NormalTok{ pos,}
    \AttributeTok{Negative =}\NormalTok{ neg,}
    \AttributeTok{Neutral =}\NormalTok{ neutral}
\NormalTok{  ))}
\NormalTok{\}}

\CommentTok{\# Apply the function to all relevant questions}
\NormalTok{question\_names }\OtherTok{\textless{}{-}} \FunctionTok{c}\NormalTok{(}\StringTok{"Q3"}\NormalTok{, }\StringTok{"Q6"}\NormalTok{, }\StringTok{"Q8"}\NormalTok{, }\StringTok{"Q10"}\NormalTok{, }\StringTok{"Q11"}\NormalTok{, }
                    \StringTok{"Q12"}\NormalTok{, }\StringTok{"Q13"}\NormalTok{, }\StringTok{"Q14"}\NormalTok{, }\StringTok{"Q15"}\NormalTok{, }\StringTok{"Q16"}\NormalTok{, }\StringTok{"Q17"}\NormalTok{)}

\ControlFlowTok{for}\NormalTok{ (qn }\ControlFlowTok{in}\NormalTok{ question\_names) \{}
  \FunctionTok{analyze\_sentiment}\NormalTok{(qn, qn)}
\NormalTok{\}}
\end{Highlighting}
\end{Shaded}

\section{Results}\label{results}

\subsection{Quantitative Results}\label{quantitative-results}

\begin{itemize}
\tightlist
\item
  Distribution of responses to Likert-scale questions
\item
  Summary statistics (mean, median, SD)
\item
  Visualizations: bar plots, histograms
\end{itemize}

\subsection{Qualitative Results}\label{qualitative-results}

The sentiment analysis revealed a clear trend of positive student
engagement across most open-ended responses. When asked about what
aspects of the course were most helpful, students consistently expressed
appreciation for the structured content delivery, use of R, and clarity
in statistical explanations. Similarly, reflections on the usefulness of
the normal model and applied examples elicited strong positive
sentiment, highlighting the effectiveness of concept integration
throughout the term.

Feedback on the clarity of teaching and accessibility of course material
was more mixed. While many students praised the instructor's effort and
teaching style, others noted inconsistencies in understanding due to the
pace of delivery or difficulty following complex R assignments.

The most critical responses were seen in prompts asking what aspects of
the course were least helpful and what improvements could be made. Here,
students frequently cited the fast pace of certain topics, confusion
around R syntax, and a need for more real-world or scaffolded examples.
Words like ``confusing,'' ``fast,'' and ``examples'' frequently appeared
in these responses, pointing to the need for clearer instructional
supports.

\begin{figure}[H]

\caption{Sentiment Distribution by Survey Question}

{\centering \includegraphics{paper_files/figure-pdf/sentiment-plot-1.pdf}

}

\end{figure}%

\section{Conclusion}\label{conclusion}

The STA107 post-course survey provides valuable insights into how
students perceive the course structure and its technical components.
Findings highlight both effective practices and opportunities for
enhancement. Continued evaluation and iteration will ensure the course
remains responsive to student needs and pedagogical best practices.

From a qualitative perspective, student feedback revealed strong
appreciation for several key elements of the course---especially the
hands-on R activities, the integration of real-world statistical
examples, and the clarity of normal model explanations. These components
were described as engaging and instructive, affirming their value in
helping students apply theoretical knowledge in practical settings.

However, when students were asked about areas of improvement, many
raised concerns regarding the speed at which complex topics were
introduced, as well as the difficulty of interpreting R-based
assignments without more structured guidance. Themes of pacing, clarity,
and the need for more scaffolded instruction were particularly
prominent. Addressing these concerns by slowing down the delivery of
technical material, providing more detailed examples, and incorporating
mid-semester feedback opportunities could significantly enhance the
learning experience in future offerings of STA107.

\section{References}\label{references}

\begin{itemize}
\item
  Jockers, M. (2017). \emph{syuzhet: Extracts Sentiment and
  Sentiment-Derived Plot Arcs from Text} (R package v1.0.4).
  \href{https://cran.r-project.org/web/packages/syuzhet/index.html}{CRAN
  Documentation}
\item
  Jockers, M. (2020). \emph{Introduction to the Syuzhet Package}.
  \href{https://cran.r-project.org/web/packages/syuzhet/vignettes/syuzhet-vignette.html}{CRAN
  Vignette}
\item
  Kim, H. (2022). Sentiment Analysis: Limits and Progress of the Syuzhet
  Package and Its Lexicons. \emph{Digital Humanities Quarterly, 16}(2).
  \href{http://www.digitalhumanities.org/dhq/vol/16/2/000601/000601.html}{Article}
\end{itemize}

\end{document}

### 2025-04-14 scripts/03-analysis_qualitative.R

@@ -1,11 +1,11 @@
 # Load required libraries
-library(tidyverse)         # For data manipulation (includes dplyr, ggplot2, etc.)
-library(stringr)           # For string handling like trimming, pattern matching
-library(tidytext)          # For text mining and tokenization
-library(syuzhet)           # For sentiment analysis using NRC lexicon
-library(ggplot2)           # For creating visualizations (e.g., bar plots)
-library(tidyr)             # For data reshaping (e.g., pivot_longer)
-library(wordcloud)         # For creating word cloud visualizations
+library(tidyverse)
+library(stringr)
+library(tidytext)
+library(syuzhet)
+library(ggplot2)
+library(tidyr)
+library(wordcloud)
 
 # Read in the survey data
 data <- read.csv("data/qualitative_analysis_data.csv", stringsAsFactors = FALSE)
@@ -26,18 +26,14 @@ sentiment_results <- data.frame(
   stringsAsFactors = FALSE
 )
 
-# Define function to calculate sentiment per question
+# Function to calculate sentiment
 analyze_sentiment <- function(question_col, question_name) {
   sentiments <- get_nrc_sentiment(data_clean[[question_col]])
   
-  # Calculate positive and negative
   pos <- sum(sentiments$positive)
   neg <- sum(sentiments$negative)
-  
-  # Define neutral as total rows - (pos > 0 or neg > 0)
   neutral <- nrow(sentiments) - sum(rowSums(sentiments[, c("positive", "negative")]) > 0)
   
-  # Append to results
   sentiment_results <<- rbind(sentiment_results, data.frame(
     Question = question_name,
     Positive = pos,
@@ -46,13 +42,23 @@ analyze_sentiment <- function(question_col, question_name) {
   ))
 }
 
-# Define the desired order of questions
-ordered_questions <- c("Q3", "Q6", "Q8", "Q10", "Q11", "Q12", "Q13", "Q14", "Q15", "Q16", "Q17")
+# Apply the function to all relevant questions
+question_names <- c("Q3", "Q6", "Q8", "Q10", "Q11", "Q12", "Q13", "Q14", "Q15", "Q16", "Q17")
 
-# Convert the 'Question' column to a factor with specified levels
-sentiment_long$Question <- factor(sentiment_long$Question, levels = ordered_questions)
+for (qn in question_names) {
+  analyze_sentiment(qn, qn)
+}
 
-# Plot with corrected order
[…]

### 2025-04-14 paper/paper.qmd
added quantitative analysis

@@
     keep-tex: true
+execute:
+  echo: false  
+  warning: false
+  message: false
 ---
@@
 qualitative_data <- read_csv("../data/qualitative_analysis_data.csv")
-# change to clean data later
+quantitative_data <- read_csv("../data/quantitative_analysis_data.csv")
 ```
@@
 
-### Qualitative Analysis
+## Quantitative Analysis
+```{r}
+# Clean column names
+colnames(quantitative_data) <- colnames(quantitative_data) %>%
+  tolower() %>%
+  str_replace_all(" ", "_") %>%
+  str_replace_all("^x", "")
+
+# Select and rename Likert-scale columns
+quantitative_data <- quantitative_data[, 1:6]
+colnames(quantitative_data) <- c(
+  "Stat_Communication_Before_Course",
+  "Stat_Communication_After_Course",
+  "R_Skills_Before_Course",
+  "R_Skills_After_Course",
+  "Understanding_After_Assignments",
+  "Understanding_After_Tutorials"
+)
+```
+
+### Communication Ability Before vs. After
+```{r}
+comm_data <- quantitative_data %>%
+  select(Stat_Communication_Before_Course, Stat_Communication_After_Course) %>%
+  pivot_longer(cols = everything(), names_to = "Question", values_to = "Rating") %>%
+  drop_na()
+
+ggplot(comm_data, aes(x = factor(Rating), fill = Question)) +
+  geom_bar(position = "dodge") +
+  theme_minimal() +
+  labs(title = "Ability to Communicate Statistics: Before vs. After",
+       x = "Rating (1 = Lowest, 5 = Highest)", y = "Count") +
+  scale_fill_manual(values = c("Stat_Communication_Before_Course" = "#66c2a5", "Stat_Communication_After_Course" = "#fc8d62"))
+```
+
+The histogram shows a clear improvement in students' perceived communication ability.
+More students rated themselves at level 4 or 5 after the course compared to before, indicating that the course helped them gain confidence in articulating statistical ideas.
+
+### R Ability Before vs. After
+```{r}
+r_data <- quantitative_data %>%
+  select(R_Skills_Before_Course, R_Skills_After_Course) %>%
+  pivot_longer(cols = everything(), names_to = "Question", values_to = "Rating") %>%
+  drop_na()
+
+ggplot(r_data, aes(x = factor(Rating), fill = Question)) +
+  geom_bar(position = "dodge") +
+  theme_minimal() +
+  labs(title = "Ability to Use R: Before vs. After",
+       x = "Rating (1 = Lowest, 5 = Highest)", y = "Count") +
+  scale_fill_manual(values = c("R_Skills_Before_Course" = "#8da0cb", "R_Skills_After_Course" = "#e78ac3"))
+```
+
+Students entered the course with relatively low self-reported R skills (mostly 1s and 2s), but after completing the course, a majority rated themselves at 3 or higher.
+This shift demonstrates the course’s success in introducing R programming effectively.
+
+### Understanding After Assignments vs. Tutorials
+```{r}
+concept_data <- quantitative_data %>%
+  select(Understanding_After_Assignments, Understanding_After_Tutorials) %>%
+  pivot_longer(cols = everything(), names_to = "Question", values_to = "Rating") %>%
+  drop_na()
+
+ggplot(concept_data, aes(x = factor(Rating), fill = Question)) +
+  geom_bar(position = "dodge") +
+  theme_minimal() +
+  labs(title = "Understanding of Statistical Concepts: Assignment vs. Tutorial",
+       x = "Rating (1 = Lowest, 5 = Highest)", y = "Count") +
+  scale_fill_manual(values = c("Understanding_After_Assignments" = "#a6d854", "Understanding_After_Tutorials" = "#ffd92f"))
+```
+
+Students found both tutorials and assignments useful for understanding statistical concepts, with a slight preference for tutorials.
+The higher number of level 4 and 5 responses for tutorials suggests that real-time, interactive support may have been especially beneficial.
+
+## Qualitative Analysis
 ```{r}
@@
 ```
-
 # Results
@@
 
-- Distribution of responses to Likert-scale questions
-- Summary statistics (mean, median, SD)
-- Visualizations: bar plots, histograms
+The survey included six Likert-scale questions aimed at capturing students’ self-perceived development in statistical understanding and communication, both before and after the course.
+The questions also addressed students’ ability to use R software and their comprehension of statistical concepts through tutorials and weekly assignments.
+
+Key findings include:
+
+- **Improved Communication Skills**: Students reported a notable improvement in their ability to communicate statistics, with the average post-course rating (M = 3.92) higher than the pre-course rating (M = 3.09).
+- **Increased R Proficiency**: The mean rating for R software usage rose from 2.43 before the course to 3.49 after completing it, indicating significant growth in computational skills.
+- **Conceptual Understanding**: Understanding of statistical concepts improved through both tutorials (M = 4.09) and weekly assignments (M = 3.84), with tutorials being rated slightly more helpful on average.
+
+The graphs above illustrate the distribution of responses across the three themes: communication, R proficiency, and statistical understanding.
 

### 2025-04-14 paper/paper.tex
Added Qual changes for themes

@@
 \makeatother
-
-\usepackage{color}
-\usepackage{fancyvrb}
-\newcommand{\VerbBar}{|}
-\newcommand{\VERB}{\Verb[commandchars=\\\{\}]}
-\DefineVerbatimEnvironment{Highlighting}{Verbatim}{commandchars=\\\{\}}
-% Add ',fontsize=\small' for more characters per line
-\usepackage{framed}
-\definecolor{shadecolor}{RGB}{241,243,245}
-\newenvironment{Shaded}{\begin{snugshade}}{\end{snugshade}}
-\newcommand{\AlertTok}[1]{\textcolor[rgb]{0.68,0.00,0.00}{#1}}
-\newcommand{\AnnotationTok}[1]{\textcolor[rgb]{0.37,0.37,0.37}{#1}}
-\newcommand{\AttributeTok}[1]{\textcolor[rgb]{0.40,0.45,0.13}{#1}}
-\newcommand{\BaseNTok}[1]{\textcolor[rgb]{0.68,0.00,0.00}{#1}}
-\newcommand{\BuiltInTok}[1]{\textcolor[rgb]{0.00,0.23,0.31}{#1}}
-\newcommand{\CharTok}[1]{\textcolor[rgb]{0.13,0.47,0.30}{#1}}
-\newcommand{\CommentTok}[1]{\textcolor[rgb]{0.37,0.37,0.37}{#1}}
-\newcommand{\CommentVarTok}[1]{\textcolor[rgb]{0.37,0.37,0.37}{\textit{#1}}}
-\newcommand{\ConstantTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{#1}}
-\newcommand{\ControlFlowTok}[1]{\textcolor[rgb]{0.00,0.23,0.31}{\textbf{#1}}}
-\newcommand{\DataTypeTok}[1]{\textcolor[rgb]{0.68,0.00,0.00}{#1}}
-\newcommand{\DecValTok}[1]{\textcolor[rgb]{0.68,0.00,0.00}{#1}}
-\newcommand{\DocumentationTok}[1]{\textcolor[rgb]{0.37,0.37,0.37}{\textit{#1}}}
-\newcommand{\ErrorTok}[1]{\textcolor[rgb]{0.68,0.00,0.00}{#1}}
-\newcommand{\ExtensionTok}[1]{\textcolor[rgb]{0.00,0.23,0.31}{#1}}
-\newcommand{\FloatTok}[1]{\textcolor[rgb]{0.68,0.00,0.00}{#1}}
-\newcommand{\FunctionTok}[1]{\textcolor[rgb]{0.28,0.35,0.67}{#1}}
-\newcommand{\ImportTok}[1]{\textcolor[rgb]{0.00,0.46,0.62}{#1}}
-\newcommand{\InformationTok}[1]{\textcolor[rgb]{0.37,0.37,0.37}{#1}}
-\newcommand{\KeywordTok}[1]{\textcolor[rgb]{0.00,0.23,0.31}{\textbf{#1}}}
-\newcommand{\NormalTok}[1]{\textcolor[rgb]{0.00,0.23,0.31}{#1}}
-\newcommand{\OperatorTok}[1]{\textcolor[rgb]{0.37,0.37,0.37}{#1}}
-\newcommand{\OtherTok}[1]{\textcolor[rgb]{0.00,0.23,0.31}{#1}}
-\newcommand{\PreprocessorTok}[1]{\textcolor[rgb]{0.68,0.00,0.00}{#1}}
-\newcommand{\RegionMarkerTok}[1]{\textcolor[rgb]{0.00,0.23,0.31}{#1}}
-\newcommand{\SpecialCharTok}[1]{\textcolor[rgb]{0.37,0.37,0.37}{#1}}
-\newcommand{\SpecialStringTok}[1]{\textcolor[rgb]{0.13,0.47,0.30}{#1}}
-\newcommand{\StringTok}[1]{\textcolor[rgb]{0.13,0.47,0.30}{#1}}
-\newcommand{\VariableTok}[1]{\textcolor[rgb]{0.07,0.07,0.07}{#1}}
-\newcommand{\VerbatimStringTok}[1]{\textcolor[rgb]{0.13,0.47,0.30}{#1}}
-\newcommand{\WarningTok}[1]{\textcolor[rgb]{0.37,0.37,0.37}{\textit{#1}}}
 
@@
 
-\subsubsection{Qualitative Analysis}\label{qualitative-analysis}
+\subsection{Quantitative Analysis}\label{quantitative-analysis}
 
-\begin{Shaded}
-\begin{Highlighting}[]
-\CommentTok{\# Load required libraries}
-\FunctionTok{library}\NormalTok{(tidyverse)}
-\FunctionTok{library}\NormalTok{(stringr)}
-\FunctionTok{library}\NormalTok{(tidytext)}
-\FunctionTok{library}\NormalTok{(syuzhet)}
-\FunctionTok{library}\NormalTok{(ggplot2)}
-\FunctionTok{library}\NormalTok{(tidyr)}
-\FunctionTok{library}\NormalTok{(wordcloud)}
-\end{Highlighting}
-\end{Shaded}
+\subsubsection{Communication Ability Before
+vs.~After}\label{communication-ability-before-vs.-after}
 
-\begin{verbatim}
-Loading required package: RColorBrewer
-\end{verbatim}
+\includegraphics{paper_files/figure-pdf/unnamed-chunk-3-1.pdf}
 
-\begin{Shaded}
-\begin{Highlighting}[]
-\CommentTok{\# Read in the survey data}
-\NormalTok{data }\OtherTok{\textless{}{-}}\NormalTok{ qualitative\_data}
+The histogram shows a clear improvement in students' perceived
+communication ability. More students rated themselves at level 4 or 5
+after the course compared to before, indicating that the course helped
+them gain confidence in articulating statistical ideas.
 
-\CommentTok{\# Rename columns}
-\FunctionTok{colnames}\NormalTok{(data)[}\DecValTok{1}\SpecialCharTok{:}\DecValTok{11}\NormalTok{] }\OtherTok{\textless{}{-}} \FunctionTok{c}\NormalTok{(}\StringTok{"Q3"}\NormalTok{, }\StringTok{"Q6"}\NormalTok{, }\StringTok{"Q8"}\NormalTok{, }\StringTok{"Q10"}\NormalTok{, }\StringTok{"Q11"}\NormalTok{, }
-                          \StringTok{"Q12"}\NormalTok{, }\StringTok{"Q13"}\NormalTok{, }\StringTok{"Q14"}\NormalTok{, }\StringTok{"Q15"}\NormalTok{, }\StringTok{"Q16"}\NormalTok{, }\StringTok{"Q17"}\NormalTok{)}
+\subsubsection{R Ability Before
+vs.~After}\label{r-ability-before-vs.-after}
 
-\CommentTok{\# Clean the text}
-\NormalTok{data\_clean }\OtherTok{\textless{}{-}}\NormalTok{ data }\SpecialCharTok{\%\textgreater{}\%}
-  \FunctionTok{mutate}\NormalTok{(}\FunctionTok{across}\NormalTok{(}\FunctionTok{everything}\NormalTok{(), }\SpecialCharTok{\textasciitilde{}} \FunctionTok{str\_trim}\NormalTok{(}\FunctionTok{tolower}\NormalTok{(}\FunctionTok{as.character}\NormalTok{(.)))))}
+\includegraphics{paper_files/figure-pdf/unnamed-chunk-4-1.pdf}
 
-\CommentTok{\# Initialize an empty data frame to collect sentiment results}
-\NormalTok{sentiment\_results }\OtherTok{\textless{}{-}} \FunctionTok{data.frame}\NormalTok{(}
-  \AttributeTok{Question =} \FunctionTok{character}\NormalTok{(),}
-  \AttributeTok{Positive =} \FunctionTok{numeric}\NormalTok{(),}
-  \AttributeTok{Negative =} \FunctionTok{numeric}\NormalTok{(),}
-  \AttributeTok{Neutral =} \FunctionTok{numeric}\NormalTok{(),}
-  \AttributeTok{stringsAsFactors =} \ConstantTok{FALSE}
-\NormalTok{)}
+Students entered the course with relatively low self-reported R skills
+(mostly 1s and 2s), but after completing the course, a majority rated
+themselves at 3 or higher. This shift demonstrates the course's success
+in introducing R programming effectively.
 
-\CommentTok{\# Function to calculate sentiment}
-\NormalTok{analyze\_sentiment }\OtherTok{\textless{}{-}} \ControlFlowTok{function}\NormalTok{(question\_col, question\_name) \{}
-\NormalTok{  sentiments }\OtherTok{\textless{}{-}} \FunctionTok{get\_nrc\_sentiment}\NormalTok{(data\_clean[[question\_col]])}
+\subsubsection{Understanding After Assignments
+vs.~Tutorials}\label{understanding-after-assignments-vs.-tutorials}
 
-\NormalTok{  pos }\OtherTok{\textless{}{-}} \FunctionTok{sum}\NormalTok{(sentiments}\SpecialCharTok{$}\NormalTok{positive)}
-\NormalTok{  neg }\OtherTok{\textless{}{-}} \FunctionTok{sum}\NormalTok{(sentiments}\SpecialCharTok{$}\NormalTok{negative)}
-\NormalTok{  neutral }\OtherTok{\textless{}{-}} \FunctionTok{nrow}\NormalTok{(sentiments) }\SpecialCharTok{{-}} 
-    \FunctionTok{sum}\NormalTok{(}\FunctionTok{rowSums}\NormalTok{(sentiments[, }\FunctionTok{c}\NormalTok{(}\StringTok{"positive"}\NormalTok{, }\StringTok{"negative"}\NormalTok{)]) }\SpecialCharTok{\textgreater{}} \DecValTok{0}\NormalTok{)}
+\includegraphics{paper_files/figure-pdf/unnamed-chunk-5-1.pdf}
 
-\NormalTok{  sentiment\_results }\OtherTok{\textless{}\textless{}{-}} \FunctionTok{rbind}\NormalTok{(sentiment\_results, }\FunctionTok{data.frame}\NormalTok{(}
-    \AttributeTok{Question =}\NormalTok{ question\_name,}
-    \AttributeTok{Positive =}\NormalTok{ pos,}
-    \AttributeTok{Negative =}\NormalTok{ neg,}
-    \AttributeTok{Neutral =}\NormalTok{ neutral}
-\NormalTok{  ))}
-\NormalTok{\}}
+Students found both tutorials and assignments useful for understanding
+statistical concepts, with a slight preference for tutorials. The higher
+number of level 4 and 5 responses for tutorials suggests that real-time,
+interactive support may have been especially beneficial.
 
-\CommentTok{\# Apply the function to all relevant questions}
-\NormalTok{question\_names }\OtherTok{\textless{}{-}} \FunctionTok{c}\NormalTok{(}\StringTok{"Q3"}\NormalTok{, }\StringTok{"Q6"}\NormalTok{, }\StringTok{"Q8"}\NormalTok{, }\StringTok{"Q10"}\NormalTok{, }\StringTok{"Q11"}\NormalTok{, }
-                    \StringTok{"Q12"}\NormalTok{, }\StringTok{"Q13"}\NormalTok{, }\StringTok{"Q14"}\NormalTok{, }\StringTok{"Q15"}\NormalTok{, }\StringTok{"Q16"}\NormalTok{, }\StringTok{"Q17"}\NormalTok{)}
-
-\ControlFlowTok{for}\NormalTok{ (qn }\ControlFlowTok{in}\NormalTok{ question\_names) \{}
-  \FunctionTok{analyze\_sentiment}\NormalTok{(qn, qn)}
-\NormalTok{\}}
-\end{Highlighting}
-\end{Shaded}
+\subsection{Qualitative Analysis}\label{qualitative-analysis}
 
@@
 
+The survey included six Likert-scale questions aimed at capturing
+students' self-perceived development in statistical understanding and
+communication, both before and after the course. The questions also
+addressed students' ability to use R software and their comprehension of
+statistical concepts through tutorials and weekly assignments.
+
+Key findings include:
+
 \begin{itemize}
@@
 \item
-  Distribution of responses to Likert-scale questions
+  \textbf{Improved Communication Skills}: Students reported a notable
+  improvement in their ability to communicate statistics, with the
+  average post-course rating (M = 3.92) higher than the pre-course
+  rating (M = 3.09).
 \item
-  Summary statistics (mean, median, SD)
+  \textbf{Increased R Proficiency}: The mean rating for R software usage
+  rose from 2.43 before the course to 3.49 after completing it,
+  indicating significant growth in computational skills.
 \item
-  Visualizations: bar plots, histograms
+  \textbf{Conceptual Understanding}: Understanding of statistical
+  concepts improved through both tutorials (M = 4.09) and weekly
+  assignments (M = 3.84), with tutorials being rated slightly more
+  helpful on average.
 \end{itemize}
+
+The graphs above illustrate the distribution of responses across the
+three themes: communication, R proficiency, and statistical
+understanding.
 

### 2025-04-14 scripts/03-analysis_qualitative.R

@@ -1,68 +1,194 @@
-# Load required libraries
-library(tidyverse)
-library(stringr)
-library(tidytext)
-library(syuzhet)
-library(ggplot2)
+#### New Code ####
+library(dplyr)
 library(tidyr)
-library(wordcloud)
+library(stringr)
+library(readr)
+library(ggplot2)
+library(tidytext)
 
-# Read in the survey data
-data <- read.csv("data/qualitative_analysis_data.csv", stringsAsFactors = FALSE)
+# 1. Read the CSV file (assumes that columns have already been cleaned and renamed)
+qual_data <- read_csv("data/cleaned_qualitative_data.csv")
 
-# Rename columns
-colnames(data)[1:11] <- c("Q3", "Q6", "Q8", "Q10", "Q11", "Q12", "Q13", "Q14", "Q15", "Q16", "Q17")
+# 2. (If needed) Rename columns – here we assume your file has columns Q3, Q6, ... already.
+#    Otherwise, you would use rename() and select(starts_with("Q")) as before.
 
-# Clean the text
-data_clean <- data %>%
-  mutate(across(everything(), ~ str_trim(tolower(as.character(.)))))
+# 3. Pivot the data into long format
+question_names <- c("Q3", "Q6", "Q8", "Q10", "Q13", "Q15", "Q16", "Q17")
+qual_long <- qual_data %>%
+  pivot_longer(cols = all_of(question_names),
+               names_to = "question",
+               values_to = "response") %>%
+  filter(!is.na(response) & response != "")
 
-# Initialize an empty data frame to collect sentiment results
-sentiment_results <- data.frame(
-  Question = character(),
-  Positive = numeric(),
-  Negative = numeric(),
-  Neutral = numeric(),
-  stringsAsFactors = FALSE
-)
-
-# Function to calculate sentiment
-analyze_sentiment <- function(question_col, question_name) {
-  sentiments <- get_nrc_sentiment(data_clean[[question_col]])
-  
-  pos <- sum(sentiments$positive)
-  neg <- sum(sentiments$negative)
-  neutral <- nrow(sentiments) - sum(rowSums(sentiments[, c("positive", "negative")]) > 0)
-  
-  sentiment_results <<- rbind(sentiment_results, data.frame(
-    Question = question_name,
-    Positive = pos,
-    Negative = neg,
-    Neutral = neutral
+# 4. Apply an extended coding scheme
+qual_coded <- qual_long %>%
+  mutate(theme = case_when(
[…]

### 2025-04-14 paper/paper.qmd
Added changes for thematic analysis

@@
 The post-course survey comprised 17 open- and close-ended questions.
-Among these, 11 were qualitative in nature (`Q3`, `Q6`, `Q8`, `Q10`, `Q11`, `Q12`, `Q13`, `Q14`, `Q15`, `Q16`, `Q17`), focusing on free-text responses reflecting students' thoughts on the content, instruction, and overall experience.
+Among these, 8 were qualitative in nature (`Q3`, `Q6`, `Q8`, `Q10`, `Q11`, `Q12`, `Q13`, `Q14`, `Q15`, `Q16`, `Q17`), focusing on free-text responses reflecting students' thoughts on the content, instruction, and overall experience.
 
@@
 # Rename columns
-colnames(data)[1:11] <- c("Q3", "Q6", "Q8", "Q10", "Q11", 
-                          "Q12", "Q13", "Q14", "Q15", "Q16", "Q17")
+colnames(data)[1:8] <- c("Q3", "Q6", "Q8", "Q10", "Q13", "Q15", "Q16", "Q17")
 
@@
 # Apply the function to all relevant questions
-question_names <- c("Q3", "Q6", "Q8", "Q10", "Q11", 
-                    "Q12", "Q13", "Q14", "Q15", "Q16", "Q17")
+question_names <- c("Q3", "Q6", "Q8", "Q10", "Q13", "Q15", "Q16", "Q17")
 
@@
 
-The sentiment analysis revealed a clear trend of positive student engagement across most open-ended responses.
-When asked about what aspects of the course were most helpful, students consistently expressed appreciation for the structured content delivery, use of R, and clarity in statistical explanations.
-Similarly, reflections on the usefulness of the normal model and applied examples elicited strong positive sentiment, highlighting the effectiveness of concept integration throughout the term.
+To gain deeper insights into student opinions, the qualitative analysis comprised two parts: thematic analysis and sentiment analysis.
 
-Feedback on the clarity of teaching and accessibility of course material was more mixed.
-While many students praised the instructor’s effort and teaching style, others noted inconsistencies in understanding due to the pace of delivery or difficulty following complex R assignments.
+### Thematic Analysis
 
-The most critical responses were seen in prompts asking what aspects of the course were least helpful and what improvements could be made.
-Here, students frequently cited the fast pace of certain topics, confusion around R syntax, and a need for more real-world or scaffolded examples.
-Words like “confusing,” “fast,” and “examples” frequently appeared in these responses, pointing to the need for clearer instructional supports.
+Following Braun and Clarke’s (2006) framework, a sample of approximately 80 student responses was examined through six phases:
+
+1. **Familiarization**: Students’ responses across 8 open-ended questions were read and annotated to gain an overall understanding.
+
+2. **Generating Initial Codes**: Semantic-level codes were created to capture features related to course experience, R skill development, and content relevance.
+
+3. **Searching for Themes**: Related codes were clustered into preliminary themes reflecting student attitudes and feedback.
+
+4. **Reviewing Themes**: Themes were refined to ensure clarity and distinctiveness across the dataset.
+
+5. **Defining and Naming Themes**: Clear definitions were established for each theme based on recurring patterns.
+
+6. **Producing the Report**: Final themes were reported below with supporting quotes.
+
+### Identified Themes
+
+#### Relevance to Real Life and Careers
+Many students expressed appreciation for statistics content when it connected to real-world examples or their field of study.
+
+> "This course gave me a solid foundation in statistical concepts and taught me how to apply them using R. Overall, I gained practical skills and became more confident in working with data."
+
+> "The course showed me how statistics applies to real-world problems, such as data analysis in my ENV course as it has alot of data."
+
+#### R as a Learning Tool
+Students had mixed responses about learning R. While some found it helpful, others found it initially intimidating.
+
+> "After the course, I had a higher opinion of my ability to use R because I learned how to create visualizations and perform statistical analyses such as normal distribution applications and regression.
+Now I can use it to organize data, generate graphs, and interpret the output.
+The tutorial session and assignments were particularly helpful in helping me gradually build my skills."
+
+> "I have never worked with R, but after trying to work with it I realized that it is also a powerful language."
+
+#### Clarity and Structure of Assignments
+Clearer instructions and expectations were a recurring theme.
+
+> "The weekly assignment was a R assignment based on the module. It was sometimes hard to understand the purpose of the task."
+
+> "After completing the weekly assignments, I have a general understanding of the statistical concepts in each module, but sometimes I still have some difficulties understanding some difficult concepts."
+
+#### Appreciation for Tutorials and TA Support
+Tutorials and TA help sessions were viewed positively, especially for R and concept reinforcement.
+
+> "For each question on the tutorial worksheet, my TA explained the relevant concepts before attempting the problem, which helped me understand the concept better.
+In addition, all tutorial exercises were related to the weekly assignment, so it gave me reassurance that I was doing everything correctly before submitting my assignments."
+
+> "The weekly tutorial was extremely useful for me.
+The worksheets were helpful and my TA, Dylan, did an excellent job at explaining the concepts involved and answering our questions when we were confused."
+
+> "my ta was great at explaining concepts and let us ask many questions which would all get responses. very good by jaiditya."
+
+#### Feedback and Reflection
+Students wanted more timely or individualized feedback.
+
+> "I always read the feedback give by the TA's since they allowed me to understand how to communicate my answers better.
+The statistical examples did help me recognize the usefulness and value of learning about statistics because it made me realize that statistics is everywhere."
+
+> "Most of the feedback I recieved was in the form of checks or X. I feel like I could have gotten a little bit better feedback but most of the assignments I did were correct."
+
+### Sentiment Analysis
+
+The sentiment analysis provided a quantitative measure of student engagement across the survey’s open-ended responses.
+Using the NRC sentiment lexicon (implemented via the syuzhet package), we quantified positive, negative, and neutral sentiments for each question.
+
+The analysis revealed:
+
+- **Predominantly Positive Sentiment**: Most responses expressed positive sentiments, especially regarding structured content delivery, effective use of R, and clear statistical explanations.
+
+- **Mixed Feedback on Clarity and Pace**: While many responses were positive overall, some negative sentiments were associated with challenges such as rapid topic progression and difficulties with R syntax.
+
+- **Critical Comments**: Negative terms like “confusing,” “fast,” and “examples” frequently appeared in responses about areas needing clearer instructional support.
+
+The bar chart above (generated in the sentiment analysis code chunk) illustrates the distribution of sentiment (positive, negative, neutral) for each of the 8 survey questions.
 
@@
 
-From a qualitative perspective, student feedback revealed strong appreciation for several key elements of the course—especially the hands-on R activities, the integration of real-world statistical examples, and the clarity of normal model explanations.
+From a qualitative perspective, student feedback revealed strong appreciation for several key elements of the course especially the hands-on R activities, the integration of real-world statistical examples, and the clarity of normal model explanations.
 These components were described as engaging and instructive, affirming their value in helping students apply theoretical knowledge in practical settings.

### 2025-04-14 paper/paper.tex
Added changes for thematic analysis

@@
 The post-course survey comprised 17 open- and close-ended questions.
-Among these, 11 were qualitative in nature (\texttt{Q3}, \texttt{Q6},
+Among these, 8 were qualitative in nature (\texttt{Q3}, \texttt{Q6},
 \texttt{Q8}, \texttt{Q10}, \texttt{Q11}, \texttt{Q12}, \texttt{Q13},
@@
 
-The sentiment analysis revealed a clear trend of positive student
-engagement across most open-ended responses. When asked about what
-aspects of the course were most helpful, students consistently expressed
-appreciation for the structured content delivery, use of R, and clarity
-in statistical explanations. Similarly, reflections on the usefulness of
-the normal model and applied examples elicited strong positive
-sentiment, highlighting the effectiveness of concept integration
-throughout the term.
+To gain deeper insights into student opinions, the qualitative analysis
+comprised two parts: thematic analysis and sentiment analysis.
 
-Feedback on the clarity of teaching and accessibility of course material
-was more mixed. While many students praised the instructor's effort and
-teaching style, others noted inconsistencies in understanding due to the
-pace of delivery or difficulty following complex R assignments.
+\subsubsection{Thematic Analysis}\label{thematic-analysis}
 
-The most critical responses were seen in prompts asking what aspects of
-the course were least helpful and what improvements could be made. Here,
-students frequently cited the fast pace of certain topics, confusion
-around R syntax, and a need for more real-world or scaffolded examples.
-Words like ``confusing,'' ``fast,'' and ``examples'' frequently appeared
-in these responses, pointing to the need for clearer instructional
-supports.
+Following Braun and Clarke's (2006) framework, a sample of approximately
+80 student responses was examined through six phases:
+
+\begin{enumerate}
+\def\labelenumi{\arabic{enumi}.}
+\item
+  \textbf{Familiarization}: Students' responses across 8 open-ended
+  questions were read and annotated to gain an overall understanding.
+\item
+  \textbf{Generating Initial Codes}: Semantic-level codes were created
+  to capture features related to course experience, R skill development,
+  and content relevance.
+\item
+  \textbf{Searching for Themes}: Related codes were clustered into
+  preliminary themes reflecting student attitudes and feedback.
+\item
+  \textbf{Reviewing Themes}: Themes were refined to ensure clarity and
+  distinctiveness across the dataset.
+\item
+  \textbf{Defining and Naming Themes}: Clear definitions were
+  established for each theme based on recurring patterns.
+\item
+  \textbf{Producing the Report}: Final themes were reported below with
+  supporting quotes.
+\end{enumerate}
+
+\subsubsection{Identified Themes}\label{identified-themes}
+
+\paragraph{Relevance to Real Life and
+Careers}\label{relevance-to-real-life-and-careers}
+
+Many students expressed appreciation for statistics content when it
+connected to real-world examples or their field of study.
+
+\begin{quote}
+``This course gave me a solid foundation in statistical concepts and
+taught me how to apply them using R. Overall, I gained practical skills
+and became more confident in working with data.''
+\end{quote}
+
+\begin{quote}
+``The course showed me how statistics applies to real-world problems,
+such as data analysis in my ENV course as it has alot of data.''
+\end{quote}
+
+\paragraph{R as a Learning Tool}\label{r-as-a-learning-tool}
+
+Students had mixed responses about learning R. While some found it
+helpful, others found it initially intimidating.
+
+\begin{quote}
+``After the course, I had a higher opinion of my ability to use R
+because I learned how to create visualizations and perform statistical
+analyses such as normal distribution applications and regression. Now I
+can use it to organize data, generate graphs, and interpret the output.
+The tutorial session and assignments were particularly helpful in
+helping me gradually build my skills.''
+\end{quote}
+
+\begin{quote}
+``I have never worked with R, but after trying to work with it I
+realized that it is also a powerful language.''
+\end{quote}
+
+\paragraph{Clarity and Structure of
+Assignments}\label{clarity-and-structure-of-assignments}
+
+Clearer instructions and expectations were a recurring theme.
+
+\begin{quote}
+``The weekly assignment was a R assignment based on the module. It was
+sometimes hard to understand the purpose of the task.''
+\end{quote}
+
+\begin{quote}
+``After completing the weekly assignments, I have a general
+understanding of the statistical concepts in each module, but sometimes
+I still have some difficulties understanding some difficult concepts.''
+\end{quote}
+
+\paragraph{Appreciation for Tutorials and TA
+Support}\label{appreciation-for-tutorials-and-ta-support}
+
+Tutorials and TA help sessions were viewed positively, especially for R
+and concept reinforcement.
+
+\begin{quote}
+``For each question on the tutorial worksheet, my TA explained the
+relevant concepts before attempting the problem, which helped me
+understand the concept better. In addition, all tutorial exercises were
+related to the weekly assignment, so it gave me reassurance that I was
+doing everything correctly before submitting my assignments.''
+\end{quote}
+
+\begin{quote}
+``The weekly tutorial was extremely useful for me. The worksheets were
+helpful and my TA, Dylan, did an excellent job at explaining the
+concepts involved and answering our questions when we were confused.''
+\end{quote}
+
+\begin{quote}
+``my ta was great at explaining concepts and let us ask many questions
+which would all get responses. very good by jaiditya.''
+\end{quote}
+
+\paragraph{Feedback and Reflection}\label{feedback-and-reflection}
+
+Students wanted more timely or individualized feedback.
+
+\begin{quote}
+``I always read the feedback give by the TA's since they allowed me to
+understand how to communicate my answers better. The statistical
+examples did help me recognize the usefulness and value of learning
+about statistics because it made me realize that statistics is
+everywhere.''
+\end{quote}
+
+\begin{quote}
+``Most of the feedback I recieved was in the form of checks or X. I feel
+like I could have gotten a little bit better feedback but most of the
+assignments I did were correct.''
+\end{quote}
+
+\subsubsection{Sentiment Analysis}\label{sentiment-analysis}
+
+The sentiment analysis provided a quantitative measure of student
+engagement across the survey's open-ended responses. Using the NRC
+sentiment lexicon (implemented via the syuzhet package), we quantified
+positive, negative, and neutral sentiments for each question.
+
+The analysis revealed:
+
+\begin{itemize}
+\item
+  \textbf{Predominantly Positive Sentiment}: Most responses expressed
+  positive sentiments, especially regarding structured content delivery,
+  effective use of R, and clear statistical explanations.
+\item
+  \textbf{Mixed Feedback on Clarity and Pace}: While many responses were
+  positive overall, some negative sentiments were associated with
+  challenges such as rapid topic progression and difficulties with R
+  syntax.
+\item
+  \textbf{Critical Comments}: Negative terms like ``confusing,''
+  ``fast,'' and ``examples'' frequently appeared in responses about
+  areas needing clearer instructional support.
+\end{itemize}
+
+The bar chart above (generated in the sentiment analysis code chunk)
+illustrates the distribution of sentiment (positive, negative, neutral)
+for each of the 8 survey questions.
 
@@
 From a qualitative perspective, student feedback revealed strong
-appreciation for several key elements of the course---especially the
+appreciation for several key elements of the course especially the
 hands-on R activities, the integration of real-world statistical

### 2025-04-14 paper/paper.qmd
Quantitative Analysis Changes made

@@
 
+\newpage
+
 # Introduction
@@
 
-This section outlines the structure of the survey and highlights key question types:
+We employed a **mixed-methods approach** that integrated both descriptive and inferential techniques for a more complete understanding of student feedback:
 
-- **Quantitative items**: Likert-scale responses on course clarity, usefulness of R exercises, and understanding of statistical concepts.
-- **Qualitative items**: Open-ended prompts for students to share what they found most/least helpful and suggestions for improvement.
+- **Quantitative analysis**: Descriptive statistics and histograms were used to examine patterns in students’ self-rated abilities before and after the course.
+In addition, **paired t-tests** were conducted to assess whether observed differences in ratings were statistically significant.
+This inferential step enhances the validity of the findings by quantifying the likelihood that the improvements were not due to chance.
 
-The post-course survey comprised 17 open- and close-ended questions.
-Among these, 8 were qualitative in nature (`Q3`, `Q6`, `Q8`, `Q10`, `Q11`, `Q12`, `Q13`, `Q14`, `Q15`, `Q16`, `Q17`), focusing on free-text responses reflecting students' thoughts on the content, instruction, and overall experience.
+- **Qualitative analysis**: Thematic coding was applied to open-ended responses to uncover recurring ideas and insights.
+We identified themes corresponding to key areas such as communication skills, R proficiency, conceptual understanding, collaboration, tutorial/TA support, and instructional tools.
+Strong supporting quotes were selected to illustrate these themes.
+Notably, many students praised the **Bookdown cherry tree assignment**, **video-based resources**, and TA assistance.
+A few comments referenced the **pace** of lectures or tutorials, but these were minimal and treated as a sub-theme.
 
-# Data Analysis
-
-## Methodology
-
-We employed a **mixed-methods approach**:
-
-- **Quantitative analysis**: Descriptive statistics and visualizations for Likert-scale questions.
-- **Qualitative analysis**: Thematic coding of open-ended responses to identify common sentiments and suggestions.
+This analysis is not a course evaluation. Rather, it reflects students’ perceptions of how the course contributed to their statistical learning and skill development across multiple dimensions.
 
@@
 ```{r}
+#| label: fig-comm-hist
+#| fig-cap: "Histogram comparing self-rated statistical communication skills before and after the course."
+
 comm_data <- quantitative_data %>%
@@
 
-The histogram shows a clear improvement in students' perceived communication ability.
+As shown in @fig-comm-hist, the histogram shows a clear improvement in students' perceived communication ability.
 More students rated themselves at level 4 or 5 after the course compared to before, indicating that the course helped them gain confidence in articulating statistical ideas.
+This shift in self-perception suggests that instruction and practice during the course likely contributed to students' ability to communicate statistical findings more effectively and with greater clarity.
+
+#### Paired t-test: Communication Skills
+
+```{r}
+#| label: tbl-comm-ttest
+
+paired_t_comm <- t.test(
+  quantitative_data$Stat_Communication_Before_Course,
+  quantitative_data$Stat_Communication_After_Course,
+  paired = TRUE
+)
+
+comm_t_table <- data.frame(
+  Mean_Before = round(mean(quantitative_data$Stat_Communication_Before_Course, 
+                           na.rm = TRUE), 2),
+  Mean_After = round(mean(quantitative_data$Stat_Communication_After_Course, 
+                          na.rm = TRUE), 2),
+  t_Statistic = round(paired_t_comm$statistic, 2),
+  p_Value = format.pval(paired_t_comm$p.value, digits = 3)
+)
+
+knitr::kable(comm_t_table, caption = "Paired t-test: Statistical Communication 
+             Skills (Before vs. After)")
+```
+
+The paired t-test in @tbl-comm-ttest shows a **statistically significant** improvement in students’ ability to communicate statistical ideas (*p* < 0.001).
 
@@
 ```{r}
+#| label: fig-r-hist
+#| fig-cap: "Histogram comparing conceptual understanding from assignments versus tutorials."
+
 r_data <- quantitative_data %>%
@@
        x = "Rating (1 = Lowest, 5 = Highest)", y = "Count") +
-  scale_fill_manual(values = c("R_Skills_Before_Course" = "#8da0cb", "R_Skills_After_Course" = "#e78ac3"))
+  scale_fill_manual(values = c("R_Skills_Before_Course" = "#8da0cb", 
+                               "R_Skills_After_Course" = "#e78ac3"))
 ```
 
-Students entered the course with relatively low self-reported R skills (mostly 1s and 2s), but after completing the course, a majority rated themselves at 3 or higher.
-This shift demonstrates the course’s success in introducing R programming effectively.
+The histogram in @fig-r-hist illustrates a noticeable increase in students' confidence using R after completing the course.
+Prior to the course, ratings clustered at the lower end (mostly 1s and 2s), suggesting limited familiarity with the software.
+After the course, a majority of responses shifted toward ratings of 3 or higher.
+This pattern implies that students gained practical experience and conceptual understanding of R through assignments, tutorials, and the hands-on structure of the course.
+
+#### Paired t-test: R Skills
+
+```{r}
+#| label: tbl-r-ttest
+
+paired_t_r <- t.test(
+  quantitative_data$R_Skills_Before_Course,
+  quantitative_data$R_Skills_After_Course,
+  paired = TRUE
+)
+
+r_t_table <- data.frame(
+  Mean_Before = round(mean(quantitative_data$R_Skills_Before_Course, na.rm = TRUE), 2),
+  Mean_After = round(mean(quantitative_data$R_Skills_After_Course, na.rm = TRUE), 2),
+  t_Statistic = round(paired_t_r$statistic, 2),
+  p_Value = format.pval(paired_t_r$p.value, digits = 3)
+)
+
+knitr::kable(r_t_table, caption = "Paired t-test: R Skills (Before vs. After)")
+```
+
+The paired t-test in @tbl-r-ttest shows a **statistically significant** improvement in students’ R programming skills (*p* < 0.001), demonstrating the course’s effectiveness in developing students' computational skills.
 
@@
 ```{r}
+#| label: fig-concept-hist
+#| fig-cap: "Histogram comparing conceptual understanding from assignments versus tutorials."
+
 concept_data <- quantitative_data %>%
@@
        x = "Rating (1 = Lowest, 5 = Highest)", y = "Count") +
-  scale_fill_manual(values = c("Understanding_After_Assignments" = "#a6d854", "Understanding_After_Tutorials" = "#ffd92f"))
+  scale_fill_manual(values = c("Understanding_After_Assignments" = "#a6d854", 
+                               "Understanding_After_Tutorials" = "#ffd92f"))
 ```
 
-Students found both tutorials and assignments useful for understanding statistical concepts, with a slight preference for tutorials.
-The higher number of level 4 and 5 responses for tutorials suggests that real-time, interactive support may have been especially beneficial.
+As shown in @fig-concept-hist, the histogram compares students’ self-rated understanding of statistical concepts based on two instructional components: assignments and tutorials.
+While both components received high ratings, tutorials showed a slight edge, especially at levels 4 and 5.
+This suggests that live, structured interaction (such as walkthroughs and direct Q&A) played a stronger role in reinforcing concepts compared to independent assignments alone.
+
+#### Paired t-test: Conceptual Understanding
+
+```{r}
+#| label: tbl-concept-ttest
+
+paired_t_concept <- t.test(
+  quantitative_data$Understanding_After_Assignments,
+  quantitative_data$Understanding_After_Tutorials,
+  paired = TRUE
+)
+
+concept_t_table <- data.frame(
+  Mean_Assignment = round(mean(quantitative_data$Understanding_After_Assignments, na.rm = TRUE), 2),
+  Mean_Tutorial = round(mean(quantitative_data$Understanding_After_Tutorials, na.rm = TRUE), 2),
+  t_Statistic = round(paired_t_concept$statistic, 2),
+  p_Value = format.pval(paired_t_concept$p.value, digits = 3)
+)
+
+knitr::kable(concept_t_table, caption = "Paired t-test: Understanding (Assignment vs. Tutorial)")
+```
+
+The t-test in @tbl-concept-ttest indicates a **statistically significant** difference favoring tutorials (*p* < 0.05), highlighting the value of interactive learning in supporting student comprehension.
 

### 2025-04-14 paper/paper.qmd
added themes for qualitatve

@@
 library(ggplot2)
+library(knitr)
 ```
@@
 }
+
+# Create question-theme mapping
+question_theme_mapping <- tibble(
+  Question = question_names,
+  Theme = c(
+    "Communication Development, Conceptual Understanding",
+    "Understanding via Tutorials & Assignments",
+    "Learning R & Technical Skills",
+    "Real-World Application",
+    "Team Collaboration",
+    "TA & Tutorial Support",
+    "Clarity of Assignments",
+    "Feedback & Reflection"
+  )
+)
 ```
@@
 The questions also addressed students’ ability to use R software and their comprehension of statistical concepts through tutorials and weekly assignments.
+
+### Summary Table of Themes per Question
 
@@
 To gain deeper insights into student opinions, the qualitative analysis comprised two parts: thematic analysis and sentiment analysis.
+
+### Summary Table of Themes per Question
+
+```{r theme-table, echo=FALSE}
+knitr::kable(theme_table, caption = "Themes Identified per Open-Ended Question")
+```
+
+### Themes and Sub-Themes
+
+#### 1. Relevance to Real Life and Careers (Q10)
+- **Sub-theme**: Connection to environmental or applied courses
+- **Frequency**: ~32 mentions
+> "The course showed me how statistics applies to real-world problems, such as data analysis in my ENV course."
+
+#### 2. R as a Learning Tool (Q8, Q6)
+- **Sub-theme**: Confidence growth
+- **Sub-theme**: Initial difficulty
+- **Frequency**: ~44 mentions
+> "After the course, I had a higher opinion of my ability to use R because I learned how to create visualizations."
+
+#### 3. Clarity and Structure of Assignments (Q16)
+- **Sub-theme**: Assignment purpose unclear
+- **Sub-theme**: Difficulty with concepts
+- **Frequency**: ~26 mentions
+> "Sometimes it was hard to understand the purpose of the task."
+
+#### 4. Appreciation for Tutorials and TA Support (Q6, Q15)
+- **Sub-theme**: TA explanation support
+- **Sub-theme**: Reinforcement through worksheets
+- **Frequency**: ~38 mentions
+> "The weekly tutorial was extremely useful. My TA explained the concepts well and answered questions."
+
+#### 5. Feedback and Reflection (Q17)
+- **Sub-theme**: Value of feedback
+- **Sub-theme**: Need for more detailed responses
+- **Frequency**: ~21 mentions
+> "Most of the feedback I received was checks or X. I would have liked more specific feedback."
 
@@
 
+This qualitative analysis reveals that students valued tutorials, TA support, and real-world connections.
+Many experienced growth in using R, while also noting areas for improvement such as assignment clarity and feedback quality.
+These insights complement the quantitative findings and can help inform future course enhancements.
+
 # Conclusion

### 2025-04-14 paper/paper.qmd
added qualitative analysis part

@@
 ```
+
+The qualitative analysis involved a combination of thematic and sentiment analysis to gain deeper insights into student experiences.
+First, student responses to open-ended survey questions were cleaned and prepared for analysis.
+Sentiment analysis was then applied using the NRC sentiment lexicon to quantify positive, negative, and neutral sentiments across each question.
+Additionally, thematic analysis was conducted to identify recurring patterns in student feedback, which were categorized into key themes such as course relevance, R proficiency, and tutorial effectiveness.
+This mixed-method approach provided a comprehensive understanding of how students felt about different aspects of the course.
+
 # Results

### 2025-04-15 paper/paper.qmd
updated pdf
final version

@@
 The paired t-test in @tbl-comm-ttest shows a **statistically significant** improvement in students’ ability to communicate statistical ideas (*p* < 0.001).
+
+\newpage
 
@@
 ```{r theme-table, echo=FALSE}
-knitr::kable(theme_table, caption = "Themes Identified per Open-Ended Question")
+knitr::kable(question_theme_mapping, caption = "Themes Identified per Open-Ended Question")
 ```
@@
 #### 3. Clarity and Structure of Assignments (Q16)
-- **Sub-theme**: Assignment purpose unclear
-- **Sub-theme**: Difficulty with concepts
-- **Frequency**: ~26 mentions
-> "Sometimes it was hard to understand the purpose of the task."
+- **Sub-theme**: Perceived challenge in task content
+- **Sub-theme**: Interpretation of assignment expectations
+- **Reflection**: Several students noted that assignments were challenging.
+This likely reflects the complex and mathematical nature of the content, an intentional design element aimed at promoting deeper learning.
+> "The assignments are not easy and you need to think critically to figure them out."
 
